---
title: "Machine Learning Week 4 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

##Data 


The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv




## Libraries
```{r,message=FALSE, warning=FALSE}
library(caret)
library(mlbench)
library(randomForest)
```

## Loading daata

```{r, cache=TRUE}
indata <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```
##Data Preparation

### Clearning up the data and Feature Selection
```{r, cache=TRUE}
##filtering columns containing only Na
indata <- indata[, colSums(is.na(testing)) != nrow(testing)]
testing <- testing[, colSums(is.na(testing)) != nrow(testing)]
##Removing unnecessary columns 
indata <- indata[,-(1:7)]

#Reducing pair-wise correlation.
correlationMatrix <- cor(indata[, -dim(indata)[2]])
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.75)
indata <- indata[, -highlyCorrelated]

```
After  removing highly correlated variables, we now have  31 features to train our models.

###Training Validation split
We split the data a 70%-30% split to training and validation respectively to get the expected out-of-sample error.
```{r, cache=TRUE}
#Spliting data into training and validation 
set.seed(123)
inTrain <- createDataPartition(y=indata$classe, p=0.7, list=F)
training <- indata[inTrain, ]
validation <- indata[-inTrain, ]
```

###Training the Model

Training a random forest model with all of the features.
```{r, cache=TRUE}
#Using Random Forest 
#Use 3-fold CV to select best  parameters
control <- trainControl(method="cv", number=3, verboseIter=F)
model_rf <- train(classe~., data=training, method="rf",  trControl=control)


model_rf$finalModel
```


###Tesing the model on the validation set.
```{r, cache=TRUE}
predictions <- predict(model_rf, newdata=validation)

confusionMatrix(validation$classe, predictions)

```

The accuracy on the validation set is 99%, thus the expected for the out-of-sample error is 1%.


###Ranking Important Features


```{r}
importance <- varImp(model_rf, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)
```

##Predicting Test set

```{r}
predictions <- predict(model_rf, testing)
predictions
```






















